Carrier phase and amplitude estimation for phase shift keying using pilots and data (Appendix)

### 2012-09-01 paper2.tex
minor edits to java options, should make the simulation go a little faster.

\begin{lemma}\label{lem:supsupDLlambda} For any $\epsilon > 0$,
\[
\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} = 0.
\]
\end{lemma}
\begin{IEEEproof}
Observe that
\begin{align*}
&\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
&=  \abs{G_L(\lambda) - T_L(\lambda) - G_L(\lambda_n) + T_L(\lambda_n)} \\
&\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
&\leq \abs{G_L(\lambda) - G_L(\lambda_n) } + \expect \abs{G_L(\lambda) - G_L(\lambda_n)},
\end{align*}
the last line following from Jensens inequality.  Put
\[
C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) },
\]
so that
\begin{align*}
V_L &= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
&\leq  C_L +  \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect\abs{G_L(\lambda) - G_L(\lambda_n) } \\
&\leq C_L + \expect C_L,
\end{align*}
where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
\[
\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
\]
Thus,
\begin{align*}
\prob&\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
\end{align*}
\end{IEEEproof}

@@

\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i, i \in D$ be defined as in Lemma~\ref{lem:diffatlambdaL}.  Then $\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i$ for all $i \in D$.
\end{lemma}
\begin{IEEEproof}
Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimsers of the function 
\[
SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
\]
defined in~\eqref{eq:SSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
\begin{equation}\label{eq:fraclambassumption}
\sfracpart{\hat{\lambda}_L + \Phi_k} > \frac{\pi}{M} + q_k
\end{equation}
for some $k \in D$.  Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.  We will show that 
\[
SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
\]
violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  First observe that,
\begin{align*}
Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*,
\end{align*}
where $\eta = e^{-j2\pi/M} - 1$ and $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
\begin{align*}
SS&(\{r_i, i \in D\}) \\
&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
&= SS(\{\hat{d}_i, i \in D\}) - D,
\end{align*}
where 
\[
D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
\]
Now $\frac{1}{L}\hat{Y} = \hat{a} = \hat{\rho} e^{j\hat{\theta}}$ from~\eqref{eq:hata} and using~\eqref{eq:yethetadhat},
\[
\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}},
\]
so that
\begin{equation}\label{eq:Ddefn}
D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
\end{equation}
Let $v = \sfracpart{\hat{\lambda}_L + \Phi_k} - \tfrac{\pi}{M}$ so that
\begin{align*}
\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) },
\end{align*}
and
\begin{align*}
\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
 &= 2 \sin(\tfrac{\pi}{M}) \sin(v),
\end{align*}
Because we assumed~\eqref{eq:fraclambassumption}, it follows that $0 > v > q_k$ and, from the definition of $q_k$,
\[
0 > \sin(v) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
\]
Subtituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  Thus,~\eqref{eq:fraclambassumption} is false.

To show that $\sfracpart{\hat{\lambda}_L + \Phi_k} \geq -\frac{\pi}{M} - q_k$ we use contradiction again.  Assume that $\sfracpart{\hat{\lambda}_L + \Phi_k} < -\frac{\pi}{M} - q_k$.  Recalling the notation $e_k^*$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k^*$.  Now an analagous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.

@@

\begin{lemma}\label{lem:Qconv}
Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
\[ 
\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where $p, d$ and $H$ are real numbers defined in the statements of Thereoms~\ref{thm:consistency}~and~\ref{thm:normality} and $o_P(1)$ denotes a number converging in probability to zero as $L\rightarrow\infty$.
\end{lemma}
\begin{IEEEproof}
We have
\[
Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0) = \tfrac{\sabs{P}}{L} k_1(\lambda) + \tfrac{\sabs{D}}{L} k_2(\lambda)
\]
where
\begin{align}
k_1(\lambda) &= \expect R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
k_2(\lambda) &= \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
\end{align}
We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\lambda = 0$,
\[
\sin(\lambda + \Phi_1) - \sin(\Phi_1) = \lambda \big( \cos(\Phi_1) + \zeta(\lambda) \big),
\]
where $\zeta$ is a continuous function with $\zeta(0) = 0$.  So,
\begin{align*}
k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
&= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big) 
\end{align*}
since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ as a result of~\eqref{eq:expectRepartRphi}.  Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e. $\zeta(\hat{\lambda}_L) = o_P(1)$.  So, 
\[
k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) \expect R_1 \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big)
\]
because $\expect R_1$ is finite.  Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
\begin{align*}
Q_L(\hat{\lambda}_L) &=  \tfrac{\sabs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\sabs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
&= \hat{\lambda}_L \left( \tfrac{\sabs{P}}{L}  + \tfrac{\sabs{D}}{L} H + o_P(1) \right) \\
&= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
\end{align*}
since $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.  The lemma follows by multiplying both sides of the above equation by $\sqrt{L}$.
\end{IEEEproof}

\begin{lemma}\label{lem:k2conv}
$k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).$
\end{lemma}
\begin{IEEEproof}
Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.  We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.  An analagous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.  To keep our notation clean put
\[
\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
\]
For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_k)$, then
\[
\fracpart{\Phi_1} = \Phi_1 - \round{\Phi_1} = \Phi_1 - \tfrac{2\pi}{M}k.
\]
Similarly, if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$, then
\[
\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k,
\]
whilst, if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$, then
\[
\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}.
\]
So, when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
\begin{align*}
\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
\end{align*}
and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
\begin{align*}
\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
&= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
\end{align*}
where, for each $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
\begin{align*}
\sin\fracpart{\lambda + \Phi_1} &- \sin\fracpart{\Phi_1} \\
&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k)
\end{align*}
and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
\begin{align*}
\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
\end{align*}
where
\[
s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
\]
and
\[
\eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos\fracpart{\Phi_1} + \zeta_{k+1}(\lambda).
\]
Observe that both $s_k$ and $\eta_k$ are continuous and that 
\[
s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
\]

>>>>

\begin{lemma}\label{lem:supsupDLlambda} For any $\epsilon > 0$,
\[
\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} = 0.
\]
\end{lemma}
\begin{IEEEproof}
Observe that
\begin{align*}
&\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
&=  \abs{G_L(\lambda) - T_L(\lambda) - G_L(\lambda_n) + T_L(\lambda_n)} \\
&\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
&\leq \abs{G_L(\lambda) - G_L(\lambda_n) } + \expect \abs{G_L(\lambda) - G_L(\lambda_n)},
\end{align*}
the last line following from Jensen's inequality.  Put
\[
C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) },
\]
so that
\begin{align*}
V_L &= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
&\leq  C_L +  \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect\abs{G_L(\lambda) - G_L(\lambda_n) } \\
&\leq C_L + \expect C_L,
\end{align*}
where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
\[
\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
\]
Thus,
\begin{align*}
\prob&\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
\end{align*}
\end{IEEEproof}

@@

\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i, i \in D$ be defined as in Lemma~\ref{lem:diffatlambdaL}.  Then $\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i$ for all $i \in D$.
\end{lemma}
\begin{IEEEproof}
Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimisers of the function 
\[
SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
\]
defined in~\eqref{eq:SSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
\begin{equation}\label{eq:fraclambassumption}
\sfracpart{\hat{\lambda}_L + \Phi_k} > \frac{\pi}{M} + q_k
\end{equation}
for some $k \in D$.  Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.  We will show that 
\[
SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
\]
violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  First observe that,
\begin{align*}
Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*,
\end{align*}
where $\eta = e^{-j2\pi/M} - 1$ and $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
\begin{align*}
SS&(\{r_i, i \in D\}) \\
&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
&= SS(\{\hat{d}_i, i \in D\}) - D,
\end{align*}
where 
\[
D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
\]
Now $\frac{1}{L}\hat{Y} = \hat{a} = \hat{\rho} e^{j\hat{\theta}}$ from~\eqref{eq:hata} and using~\eqref{eq:yethetadhat},
\[
\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}},
\]
so that
\begin{equation}\label{eq:Ddefn}
D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
\end{equation}
Let $v = \sfracpart{\hat{\lambda}_L + \Phi_k} - \tfrac{\pi}{M}$ so that
\begin{align*}
\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) },
\end{align*}
and
\begin{align*}
\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
 &= 2 \sin(\tfrac{\pi}{M}) \sin(v),
\end{align*}
Because we assumed~\eqref{eq:fraclambassumption}, it follows that $0 > v > q_k$ and, from the definition of $q_k$,
\[
0 > \sin(v) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
\]
Substituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  Thus,~\eqref{eq:fraclambassumption} is false.

To show that $\sfracpart{\hat{\lambda}_L + \Phi_k} \geq -\frac{\pi}{M} - q_k$ we use contradiction again.  Assume that $\sfracpart{\hat{\lambda}_L + \Phi_k} < -\frac{\pi}{M} - q_k$.  Recalling the notation $e_k^*$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k^*$.  Now an analogous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.

@@

\begin{lemma}\label{lem:Qconv}
Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
\[ 
\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where $p, d$ and $H$ are real numbers defined in the statements of Theorems~\ref{thm:consistency}~and~\ref{thm:normality} and $o_P(1)$ denotes a number converging in probability to zero as $L\rightarrow\infty$.
\end{lemma}
\begin{IEEEproof}
We have
\[
Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0) = \tfrac{\sabs{P}}{L} k_1(\lambda) + \tfrac{\sabs{D}}{L} k_2(\lambda)
\]
where
\begin{align}
k_1(\lambda) &= \expect R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
k_2(\lambda) &= \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
\end{align}
We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\lambda = 0$,
\[
\sin(\lambda + \Phi_1) - \sin(\Phi_1) = \lambda \big( \cos(\Phi_1) + \zeta(\lambda) \big),
\]
where $\zeta$ is a continuous function with $\zeta(0) = 0$.  So,
\begin{align*}
k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
&= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big) 
\end{align*}
since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ as a result of~\eqref{eq:expectRepartRphi}.  Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e. $\zeta(\hat{\lambda}_L) = o_P(1)$.  So, 
\[
k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) \expect R_1 \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big)
\]
because $\expect R_1$ is finite.  Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
\begin{align*}
Q_L(\hat{\lambda}_L) &=  \tfrac{\sabs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\sabs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
&= \hat{\lambda}_L \left( \tfrac{\sabs{P}}{L}  + \tfrac{\sabs{D}}{L} H + o_P(1) \right) \\
&= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
\end{align*}
since $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.  The lemma follows by multiplying both sides of the above equation by $\sqrt{L}$.
\end{IEEEproof}

\begin{lemma}\label{lem:k2conv}
$k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).$
\end{lemma}
\begin{IEEEproof}
Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.  We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.  An analogous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.  To keep our notation clean put
\[
\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
\]
For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_k)$, then
\[
\fracpart{\Phi_1} = \Phi_1 - \round{\Phi_1} = \Phi_1 - \tfrac{2\pi}{M}k.
\]
Similarly, if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$, then
\[
\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k,
\]
whilst, if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$, then
\[
\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}.
\]
So, when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
\begin{align*}
\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
\end{align*}
and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
\begin{align*}
\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
&= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
\end{align*}
where, for each $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
\begin{align*}
\sin\fracpart{\lambda + \Phi_1} &- \sin\fracpart{\Phi_1} \\
&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k)
\end{align*}
and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
\begin{align*}
\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
\end{align*}
where
\[
s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
\]
and
\[
\eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos\fracpart{\Phi_1} + \zeta_{k+1}(\lambda).
\]
Observe that both $s_k$ and $\eta_k$ are continuous and that 
\[
s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
\]

Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
\[
\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
\]
whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
\[
\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
\]
Since the intervals $[\psi_{k-1}, \psi_{k})$ partition the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
\[
\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
\]
where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
\begin{equation}\label{eq:vk}
v_k(\lambda,\Phi_1) = \begin{cases}
s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
\lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
0 & \text{otherwise}.
\end{cases} 
\end{equation}
Now,
\begin{align*}
k_2(\lambda) &=  \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
&= \expect R_1\big( \lambda\cos\fracpart{\Phi_1} + v(\lambda, \Phi_1)  \big) \\
&= \lambda h_2(0) + \expect R_1 v(\lambda, \Phi_1).
\end{align*}
Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a continuous function with $\gamma(0) = 0$, and
\[
H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0),
\]
where $H$ is defined in the statement of Theorem~\ref{thm:normality}.  Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, it follows that $\gamma(\hat{\lambda}_L) = o_P(1)$, and so
\[
k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( h_2(0) + H_1 + o_P(1) ) = \hat{\lambda}_L ( H + o_P(1) )
\]
as required.  It remains to prove Lemma~\ref{lem:expRvlamphi}.
\end{IEEEproof}
